\documentclass[10pt]{article}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage{url}
\usepackage{multicol}
\usepackage{xspace}
\usepackage{pstricks}
\usepackage{wrapfig}
\usepackage[section]{placeins}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage[default]{droidserif}
\usepackage[T1]{fontenc}

\usepackage{listings}

%\usepackage{algorithm2e}
\usetikzlibrary{arrows,automata,shapes}
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=2.5em, text centered, rounded corners, minimum height=2em]
\tikzstyle{bw} = [rectangle, draw, fill=blue!20, 
    text width=4em, text centered, rounded corners, minimum height=2em]

\newcommand{\handout}[5]{
  \noindent
  \begin{center}
  \framebox{
    \vbox{
      \hbox to 5.78in { {\bf ECE155: Engineering Design with Embedded Systems } \hfill #2 }
      \vspace{4mm}
      \hbox to 5.78in { {\Large \hfill #5  \hfill} }
      \vspace{2mm}
      \hbox to 5.78in { {\em #3 \hfill #4} }
    }
  }
  \end{center}
  \vspace*{4mm}
}

\newcommand{\lecture}[4]{\handout{#1}{#2}{#3}{#4}{Lab #1}}
\topmargin 0pt
\advance \topmargin by -\headheight
\advance \topmargin by -\headsep
\textheight 8.9in
\oddsidemargin 0pt
\evensidemargin \oddsidemargin
\marginparwidth 0.5in
\textwidth 6.5in

\parindent 0in
\parskip 1.5ex
%\renewcommand{\baselinestretch}{1.25}

\newcommand{\todo}[1]{{\red\textbf{TODO: }#1}\xspace}

\lstset{language=java, 
        basicstyle=\ttfamily,columns=fullflexible,
%	keywordstyle=\color{Blue},          % keyword style
%	commentstyle=\color{OliveGreen}\textit,       % comment style
%	identifierstyle=\color{Black},
%	stringstyle=\color{BrickRed},
	mathescape=true,
	tabsize=3,
	showstringspaces=false}

\begin{document}

\lecture{2 (Detecting Footsteps) --- Assessment}{Winter 2015}{Prepared by Kirill Morozov}{version 1.1}

You are responsible for conforming to ``University of Waterloo Policy 71: Student Academic Discipline.''  Students complete Part I of this form.  The TA conducting the demo completes the rest after the demo. We are entering marks based on this form, so if no form exists, you get no marks.

\vspace{-1em}

\section*{Part I: Student Comments}
The design (check one of the following): 
\begin{itemize}
\renewcommand{\labelitemi}{$\Box$}
\item Does not incorporate others' work with the exception of the university-provided materials.
\item Incorporates the work of others as indicated in the notes below. 
\end{itemize}
By signing below, I confirm that we wrote the submitted lab code and that it has not been previously submitted for academic credit at this or any other academic institution, except as noted below.

\begin{tabular*}{1\textwidth}{@{\extracolsep{\fill} }c|c|c|c|}
\hline
&Student Name & UW Student ID \# & Signature \\
\hline
Student 1&&&\\[1em]
\hline
Student 2&&&\\[1em]
\hline
Student 3 &&&\\[1em]
\hline
\end{tabular*}

\paragraph{Notes:}


\newpage
\section*{Part II: Demonstration Checklist}
\paragraph{Software Design Checklist (1 mark for each checklist item satisfied)}
\begin{itemize}
\renewcommand{\labelitemi}{$\Box$}
\item Deductions Only: Solution was committed to SVN and compiles without errors \\
(This item is mandatory; you get a 0 for broken or uncommitted solutions.)
\item The number of steps your application has counted is clearly displayed.
\item There is a way inside the application to reset the step counter.
\item The group members can explain how their step-detection algorithm works.
\renewcommand{\labelitemi}{$\Box\Box$}
\item Your error rate, both false positives and false negatives, is less than once in every three steps
(33\% error).
\item Your error rate, both false positives and false negatives, is less than once in every ten steps
(10\% error)
\item The design and implementation follow good engineering design. Examples: not over-using global variables, avoiding unnecessary code duplication, and giving variables descriptive names.
\renewcommand{\labelitemi}{$\Box$}
\item You have eliminated one of the sources of error mentioned in the lab manual and can explain your solution to the TA.
\end{itemize}

\newpage
\section*{Part III: TA Comments}

\paragraph{Lab \#:}
\paragraph{Group \#:}
\paragraph{Date and Time:}

\paragraph{TA Name:}
\paragraph{Signature:}

\paragraph{Mark:}
\paragraph{Raw Trial Data:} ~\\[1em]
\hspace*{2em} \begin{tabular*}{0.75\textwidth}{@{\extracolsep{\fill} }c|c|c|c|c|c}
&Trial 1 & Trial 2& Trial 3 & Trial 4 & Trial 5 \\
\hline
Steps Taken &&&&&\\[1em]
\hline
Steps Counted &&&&&\\[1em]
\end{tabular*}

\paragraph{Notes:}

\end{document}

