\documentclass[10pt]{article}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage{url}
\usepackage{multicol}
\usepackage{xspace}
\usepackage{pstricks}
\usepackage{wrapfig}
\usepackage[section]{placeins}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage[default]{droidserif}
\usepackage[T1]{fontenc}

\usepackage{listings}

%\usepackage{algorithm2e}
\usetikzlibrary{arrows,automata,shapes}
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=2.5em, text centered, rounded corners, minimum height=2em]
\tikzstyle{bw} = [rectangle, draw, fill=blue!20, 
    text width=4em, text centered, rounded corners, minimum height=2em]

\newcommand{\handout}[5]{
  \noindent
  \begin{center}
  \framebox{
    \vbox{
      \hbox to 5.78in { {\bf ECE155: Engineering Design with Embedded Systems } \hfill #2 }
      \vspace{4mm}
      \hbox to 5.78in { {\Large \hfill #5  \hfill} }
      \vspace{2mm}
      \hbox to 5.78in { {\em #3 \hfill #4} }
    }
  }
  \end{center}
  \vspace*{4mm}
}

\newcommand{\lecture}[4]{\handout{#1}{#2}{#3}{#4}{Lab #1}}
\topmargin 0pt
\advance \topmargin by -\headheight
\advance \topmargin by -\headsep
\textheight 8.9in
\oddsidemargin 0pt
\evensidemargin \oddsidemargin
\marginparwidth 0.5in
\textwidth 6.5in

\parindent 0in
\parskip 1.5ex
%\renewcommand{\baselinestretch}{1.25}

\newcommand{\todo}[1]{{\red\textbf{TODO: }#1}\xspace}

\lstset{language=java, 
        basicstyle=\ttfamily,columns=fullflexible,
%	keywordstyle=\color{Blue},          % keyword style
%	commentstyle=\color{OliveGreen}\textit,       % comment style
%	identifierstyle=\color{Black},
%	stringstyle=\color{BrickRed},
	mathescape=true,
	tabsize=3,
	showstringspaces=false}

\begin{document}

\lecture{3 (Dead Reckoning) --- Assessment}{Winter 2015}{Prepared by Kirill Morozov}{version 1.1}


You are responsible for conforming to ``University of Waterloo Policy 71: Student Academic Discipline.''  Students complete Part I of this form.  The TA conducting the demo completes the rest after the demo. We are entering marks based on this form, so if no form exists, you get no marks.

\vspace{-1em}

\section*{Part I: Student Comments}
The design (check one of the following): 
\begin{itemize}
\renewcommand{\labelitemi}{$\Box$}
\item Does not incorporate others' work with the exception of the university-provided materials.
\item Incorporates the work of others as indicated in the notes below. 
\end{itemize}
By signing below, I confirm that we wrote the submitted lab code and that it has not been previously submitted for academic credit at this or any other academic institution, except as noted below.

\begin{tabular*}{1\textwidth}{@{\extracolsep{\fill} }c|c|c|c|}
\hline
&Student Name & UW Student ID \# & Signature \\
\hline
Student 1&&&\\[1em]
\hline
Student 2&&&\\[1em]
\hline
Student 3 &&&\\[1em]
\hline
\end{tabular*}

\paragraph{Notes:}


\newpage
\section*{Part II: Demonstration Checklist}

Calculating displacement error procedure:  Since the net displacement displayed on the screen is in North-South and East-West components, the TA will do the vector summation to get the total steps miscounted, and divide this number against the total steps taken to get the percentage error rate.

TAs can ask the you to repeat the evaluations up to 5 times, and the best 3 runs will be the ones evaluated, provided that the algorithm you have developed is reasonably functional.

\paragraph{Software Design Checklist (1 mark for each checklist item satisfied)}
\begin{itemize}
\renewcommand{\labelitemi}{$\Box$}
\item Solution was committed to SVN and compiles without errors \\
(This item is mandatory; you get a 0 for broken or uncommitted solutions.)
\item The number of steps your application has counted in two orthogonal directions is clearly displayed.
\item There is a way inside the application to reset the step counter.
\item You have loaded the map and show it correctly in the screen.
\item North-South Calibration can be conducted.
\renewcommand{\labelitemi}{$\Box\Box\Box\Box\Box\Box$}
\item Your error rate, in false positives and negatives, is less than 25\%. An additional 0.5pt for each 5\% increase accuracy; if accuracy is 70\% or below, no points (3 pts max if 100\% accurate).
\renewcommand{\labelitemi}{$\Box\Box$}
\item The design and implementation follow good engineering design. Examples: not over-using global variables, avoiding unnecessary code duplication, and giving variables descriptive names.
\end{itemize}


\newpage
\newpage
\section*{Part III: TA Comments}

\paragraph{Lab \#:}
\paragraph{Group \#:}
\paragraph{Date and Time:}

\paragraph{TA Name:}
\paragraph{Signature:}

\paragraph{Mark:}
\paragraph{Raw Trial Data:} ~\\[1em]
\hspace*{2em} \begin{tabular*}{0.75\textwidth}{@{\extracolsep{\fill} }c|c|c|c}
&Trial 1 & Trial 2& Trial 3 \\
\hline
Steps Taken &&&\\[1em]
\hline
Steps Counted &&&\\[1em]
\end{tabular*}

\paragraph{Notes:}


\end{document}

